\documentclass[aspectratio=169,11pt]{beamer}

\usetheme{metropolis}
\usecolortheme{default}

% \usetheme{Boadilla}
% \usecolortheme{orchid}
% \usecolortheme{dove}
% \usecolortheme{beaver}
% \usecolortheme{whale}



\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}{%
  \hfill\usebeamerfont{page number in head/foot}\insertframenumber\hspace{3mm}\vspace{2mm}
}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}
\usepackage{array}
\usepackage{xcolor}
\usepackage{appendixnumberbeamer}

\usetikzlibrary{arrows.meta,positioning,calc,fit,shapes.geometric,decorations.pathreplacing}
\pgfplotsset{compat=1.18}

\definecolor{sbbred}{HTML}{E0001B}
\definecolor{ink}{HTML}{202124}
\definecolor{muted}{HTML}{5F6368}
\definecolor{track}{HTML}{2E5EAA}
\definecolor{walk}{HTML}{D9822B}
\definecolor{okgreen}{HTML}{16803C}
\definecolor{warn}{HTML}{B45309}
\definecolor{softblue}{HTML}{EAF2FF}
\definecolor{softgreen}{HTML}{EAF7EF}
\definecolor{softorange}{HTML}{FFF2E0}
\definecolor{softred}{HTML}{FDE8EA}

\setbeamercolor{normal text}{fg=ink,bg=white}
\setbeamercolor{frametitle}{fg=ink,bg=white}
\setbeamercolor{progress bar}{fg=sbbred,bg=softred}
\setbeamercolor{alerted text}{fg=sbbred}

\newcommand{\ph}[1]{\textcolor{warn}{\textbf{PLACEHOLDER: #1}}}
\newcommand{\smallcap}[1]{\textcolor{muted}{\scriptsize\MakeUppercase{#1}}}
\newcommand{\metric}[2]{%
  \begin{tabular}{@{}c@{}}
    {\Large\bfseries #1}\\[-1mm]
    {\scriptsize\textcolor{muted}{#2}}
  \end{tabular}}

\title{Robust Journey Planning for SBB}
\subtitle{Fastest route that still arrives on time with probability at least $Q$}
\author{Zhixiang Dai, Weijiang Xiong, Jinghao Zheng, Kaile Chu}
% \author{Z. Dai, W. Xiong, J. Zheng, K. Chu}
\date{May 27, 2026}

\begin{document}

% ---------------------------------------------------------------------------
\begin{frame}
  \titlepage
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{The Question We Answer}

  \vspace{-3mm}
  \begin{columns}[T,onlytextwidth]
    \column{0.49\textwidth}
      \begin{block}{Input}
        \begin{itemize}
          \item origin stop $A$, destination stop $B$
          \item weekday and latest arrival time $T$
          \item confidence threshold $Q$
          \item region UUID list from \texttt{geo}
        \end{itemize}
      \end{block}

    \vspace{-2.5mm}  
      \begin{block}{Output}
      \begin{itemize}
          \item Routes arriving before $T$ with confidence $\ge Q$, sorted by latest departure time.
      \end{itemize}
      \end{block}

    \column{0.49\textwidth}
      \input{figures/concept/question_routes.tex}
  \end{columns}

  \begin{alertblock}{Main assumptions}
    \vspace{1mm}
    Latest timetable week; no holiday exceptions; walking is deterministic at 50 m/min; route leg delays are combined under an independence assumption.
  \end{alertblock}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{System Overview}
  \centering

    \resizebox{\textwidth}{!}{\input{figures/system/overview.tex}}

  \vspace{2mm}
  \only<1>{\smallcap{Build a planner from timetable, history, and geography.}}
  \only<2>{\smallcap{Preparation builds reusable regional tables and delay distributions.}}
  \only<3>{\smallcap{Each user query runs routing plus confidence scoring.}}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{Data Structure: A Time-Dependent Transit Graph}
  \centering
  \resizebox{0.95\textwidth}{!}{\input{figures/model/time_dependent_graph.tex}}

  \vspace{-5mm}
  \begin{block}{Implementation detail}
    \small\texttt{prepare(regions=[...])} rebuilds regional \texttt{stops}, \texttt{stop\_to\_stop}, and \texttt{stop\_times}.
  \end{block}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{Delay Model: Probability From Historical Slack}
  \begin{columns}[T,onlytextwidth]
    \column{0.48\textwidth}
      \begin{block}{Training signal}
        \small
        \[
          \mathrm{delay}
          = t_{\mathrm{actual\ arrival}}
          - t_{\mathrm{scheduled\ arrival}}
        \]
      \end{block}

      Clean Istdaten records:
      \begin{itemize}
        \item planned and actual arrival known
        \item not failed, not unplanned
        \item arrival status marked as real
      \end{itemize}

      \vspace{1mm}
      Profiles are grouped by:
      \[
        \text{stop} \times \text{weekday} \times \text{scheduled\_arrival\_hour}
      \]

    \column{0.48\textwidth}
      \centering
      \input{figures/model/delay_slack_profile.tex}

      \vspace{-1mm}
      {\scriptsize\textcolor{muted}{QUANTILE PROFILES SCORE ANY SLACK VALUE.}}
  \end{columns}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{How Confidence Enters a Route}
  \centering
  \input{figures/model/route_confidence.tex}

  \vspace{1mm}
  \only<1>{\smallcap{Each transfer/final deadline creates a slack value.}}
  \only<2>{\smallcap{The project allows assuming independent delay events.}}
  \only<3>{\smallcap{Reliability filters routes before ranking.}}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{Routing Algorithm}
  \begin{columns}[T,onlytextwidth]
    \column{0.66\textwidth}
      \begin{block}{Candidate generation}
        \begin{enumerate}
          \item Sweep departure probes backward from $T$.
          \item Run Multi-Criteria RAPTOR for each probe.
          \item Keep Pareto journeys by arrival time, walking distance, and transfers.
          \item Score each route with the empirical confidence model.
        \end{enumerate}
      \end{block}

      \vspace{1mm}
      \begin{alertblock}{Ranking contract}
        Return routes with confidence $\ge Q$, sorted from latest departure to earliest departure.
      \end{alertblock}

    \column{0.29\textwidth}
      \input{figures/routing/algorithm_pipeline.tex}
  \end{columns}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{Validation: Does the Probability Mean What It Says?}
  \begin{columns}[T,onlytextwidth]
    \column{0.46\textwidth}
      \begin{block}{Strict time split}
        \begin{itemize}
          \item Training: max date $-365$ days to max date $-30$ days
          \item Test: final 30 days, never used in training
          \item No future information is used
        \end{itemize}
      \end{block}

      \begin{block}{Metrics}
        \small
        \begin{itemize}
          \item calibration error at p50, p80, p90, p95
          \item per-stop calibration error
          \item robust planner vs. fastest-route baseline
        \end{itemize}
      \end{block}

    \column{0.50\textwidth}
      \centering
      \input{figures/validation/calibration_plot.tex}

      \vspace{-2mm}
      \ph{Insert final calibration values from cluster run.}
  \end{columns}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{Validation Finding: Morning Peak Risk}
  \begin{columns}[T,onlytextwidth]
    \column{0.53\textwidth}
      \centering
      \input{figures/validation/morning_peak_risk.tex}

      \ph{Replace with final hourly p90 chart if rerun.}

    \column{0.43\textwidth}
      \begin{block}{Interpretation}
        The empirical model is broadly calibrated, but morning-peak cascade delays are harder to capture.
      \end{block}

      \begin{itemize}
        \item Hour 7 is the known weak point in the current notebook.
        \item This is statistically meaningful when the test count is large.
        \item For 07:00 journeys, the planner may over-promise unless extra buffer is required.
      \end{itemize}

      \vspace{1mm}
      \smallcap{This limitation is part of the method, not hidden from the user.}
  \end{columns}
\end{frame}

% ---------------------------------------------------------------------------

\begin{frame}{Route-Based Validation Results}
  \begin{columns}[T,onlytextwidth]
    \column{0.65\textwidth}
      \begin{itemize}
        \item Reserve 30 days of data to validate the planned routes, delay model trained without these days.
        \item Travel Demand: random pairs of origin-destination stops, random day of week, random arrival deadline from 7:00 to 22:00 with 30-minute intervals.
        \item Match with istdaten records by stop ID and scheduled arrival time
        \item Pure-walking routes or those with unmatched records are excluded before validation.
      \end{itemize}

    \column{0.35\textwidth}
      \centering
      \begin{tabular}{lr}
        \multicolumn{2}{l}{\bfseries Validation summary}\\[1mm]
        \toprule
        Confidence & 90.0\%\\
        Routes & 50\\
        Validation & 195\\
        Successful & 162 (83.1\%)\\
        \bottomrule
      \end{tabular}

      \vspace{3mm}
      \begin{tabular}{lrr}
        \multicolumn{3}{l}{\bfseries Per-Route Success}\\[1mm]
        \toprule
        rate & routes & share\\
        \midrule
        100\% & 35 & 70\%\\
        75\% & 6 & 12\%\\
        $\leq$50\% & 9 & 18\%\\
        \bottomrule
      \end{tabular}
  \end{columns}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{Operational Example}
  \begin{columns}[T,onlytextwidth]
    \column{0.45\textwidth}
      \begin{block}{Query}
        \begin{tabular}{@{}ll@{}}
          Origin & EPFL (8501214)\\
          Destination & CHUV (8592017)\\
          Day & Monday\\
          Arrival deadline & 09:00\\
          Confidence & 90\%\\
          Walking budget & 500 m
        \end{tabular}
      \end{block}

      \begin{block}{Result}
        \centering
        \begin{tabular}{ccc}
          \toprule
          dep. & arr. & confidence\\
          \midrule
          07:47 & 08:08 & 96.6\%\\
          07:42 & 08:04 & 95.0\%\\
          \bottomrule
        \end{tabular}
      \end{block}

    \column{0.51\textwidth}
      \centering
      \resizebox{\linewidth}{!}{%
      \input{figures/examples/operational_route_sketch.tex}
      }

      \vspace{2mm}
      \smallcap{No walking leg; transfer slack at Lausanne-Flon is 4 min / 3 min.}
  \end{columns}
\end{frame}

% ---------------------------------------------------------------------------
\begin{frame}{What We Can Claim}
  \small
  \begin{columns}[T,onlytextwidth]
    \column{0.50\textwidth}
      \begin{block}{Strengths}
        \begin{itemize}
          \item Region-agnostic initialization via geo UUIDs.
          \item Explainable delay model tied directly to route slack.
          \item Standard public-transit algorithm adapted to walking and reliability constraints.
          \item Validation uses held-out future days, avoiding leakage.
        \end{itemize}
      \end{block}

    \column{0.46\textwidth}
      \begin{alertblock}{Limitations}
        \begin{itemize}
          \item Independence assumption between leg delays.
          \item Walking is straight-line and deterministic.
          \item McRAPTOR candidates are Pareto by arrival/walking/transfers before confidence scoring.
          \item Morning-peak reliability is currently less well calibrated.
        \end{itemize}
      \end{alertblock}
  \end{columns}

  \vspace{2mm}
  \begin{center}
    \normalsize\bfseries
    The planner returns the fastest route that is reliable enough.
  \end{center}
\end{frame}

% ---------------------------------------------------------------------------
\appendix
\begin{frame}{Backup: Suggested Final Numbers To Fill In}
  \begin{itemize}
    \item Region names and UUIDs used in the video.
    \item Counts after preprocessing: stops, walking edges, stop-time rows, grouped routes.
    \item Delay-profile rows and training/test date range.
    \item Calibration table: target quantile, observed fraction, Wilson CI, error.
    \item Per-stop calibration error vs. global baseline.
    \item One route example: departure, arrival, confidence, transfer slack.
    \item Runtime: preparation time and one-query routing time.
  \end{itemize}
\end{frame}

\end{document}
